# AOL LAB Data Mining and Visualization

#### Shafa Amira Qonitatin - 2502009173

## About Dataset

-   gender - Whether the customer is a male or a female
-   SeniorCitizen - Whether the customer is a senior citizen or not (1, 0)
-   Partner - Whether the customer has a partner or not (Yes, No)
-   Dependents - Whether the customer has dependents or not (Yes, No)
-   Tenure - Number of months the customer has stayed with the company check
-   PhoneService - Whether the customer has a phone service or not (Yes, No)
-   MultipleLines - Whether the customer has multiple lines or not (Yes, No, No phone service)
-   InternetService - Customer's internet service provider (DSL, Fiber optic, No)
-   OnlineSecurity - Whether the customer has online security or not (Yes, No, No internet service)
-   OnlineBackup - Whether the customer has online backup or not (Yes, No, No internet service)
-   DeviceProtection - Whether the customer has device protection or not (Yes, No, No internet service)
-   TechSupport - Whether the customer has tech support or not (Yes, No, No internet service)
-   StreamingTV - Whether the customer has streaming TV or not (Yes, No, No internet service)
-   StreamingMovies - Whether the customer has streaming movies or not (Yes, No, No internet service)
-   Contract - The contract term of the customer (Month-to-month, One year, Two year)
-   PaperlessBilling - Whether the customer has paperless billing or not (Yes, No)
-   PaymentMethod - The customer's payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit card (automatic))
-   MonthlyCharges - The amount charged to the customer monthly
-   TotalCharges - The total amount charged to the customer
-   Churn - Whether the customer churned or not (Yes or No)

Dataset is from this following link: <https://www.kaggle.com/datasets/blastchar/telco-customer-churn>

# Exploratory Data Analysis (EDA)

## 1) Understanding Variables

```{r}
# load dataset
library(readr)
df = read_csv("/Users/shafaqonitatingmail.com/Documents/Semester 4/Data Mining and Visualization/Assignment/LAB/Telco Customer Churn/WA_Fn-UseC_-Telco-Customer-Churn.csv", show_col_types = FALSE)
```

```{r}
# observation the first 6 column
head(df)
```

```{r}
# view the structure of the dataset
str(df)
```

```{r}
# view summary statistics
summary(df)
```

```{r}
# misformatted data
print(unique(df$gender))
print(unique(df$Partner))
print(unique(df$Dependents))
print(unique(df$PhoneService))
print(unique(df$MultipleLines))
print(unique(df$InternetService))
print(unique(df$OnlineSecurity))
print(unique(df$OnlineBackup))
print(unique(df$DeviceProtection))
print(unique(df$TechSupport))
print(unique(df$StreamingTV))
print(unique(df$StreamingMovies))
print(unique(df$Contract))
print(unique(df$PaperlessBilling))
print(unique(df$PaymentMethod))
print(unique(df$Churn))
```

## 2) Cleaning dataset

```{r}
# delete customerID
df = subset(df,select = -c(customerID))
```

```{r}
head(df)
```

```{r}
df$SeniorCitizen <- ifelse(df$SeniorCitizen == 1, "Yes", "No")
```

```{r}
head(df)
```

```{r}
# omit null values
df <- na.omit(df)
```

```{r}
summary(df)
```

```{r}
library(ggplot2)

# outliers
b1 = ggplot(data = df, aes(x = "", y = tenure)) +
  geom_boxplot(fill = "lightblue", color = "#142850") +
  labs(x = "", y = "Tenure", title = "Boxplot of Tenure")

b2 = ggplot(data = df, aes(x = "", y = MonthlyCharges)) +
  geom_boxplot(fill = "lightblue", color = "#142850") +
  labs(x = "", y = "MonthlyCharges", title = "Boxplot of Monthly Charges")

b3 = ggplot(data = df, aes(x = "", y = TotalCharges)) +
  geom_boxplot(fill = "lightblue", color = "#142850") +
  labs(x = "", y = "TotalCharges", title = "Boxplot of Total Charges")

b1
b2
b3
```

## 3) Visualization

```{r}
# Histogram Plot Distribution of tenure
h1 =ggplot(data = df, aes(x = tenure)) +
  geom_histogram(fill = "lightblue", color = "#142850", bins = 30) +
  labs(x = "Tenure", y = "Count of Tenure", title = "Histogram Plot Distribution of Tenure")

# Histogram Plot Distribution of MonthlyCharges
h2 = ggplot(data = df, aes(x = MonthlyCharges)) +
  geom_histogram(fill = "lightblue", color = "#142850", bins = 30) +
  labs(x = "MonthlyCharges", y = "Count of MonthlyCharges", title = "Histogram Plot Distribution of MonthlyCharges")

# Histogram Plot Distribution of TotalCharges
h3 = ggplot(data = df, aes(x = TotalCharges)) +
  geom_histogram(fill = "lightblue", color = "#142850", bins = 30) +
  labs(x = "TotalCharges", y = "Count of TotalCharges", title = "Histogram Plot Distribution of TotalCharges")

h1
h2
h3
```

```{r}
ggplot(data = df, aes(x = Churn)) +
  geom_bar(fill = c("#142850", "lightblue")) +
  labs(x = "Churn", y = "Count", title = "Bar Plot Distribution of Churn")
```

```{r}
# Contract Distribution Bar Plot with Churn Grouped
g1 <- ggplot(data = df, aes(x = Contract)) +
  geom_bar(aes(fill = Churn), position = "dodge") +
  labs(x = "Contract", y = "Count", title = "Contract Distribution Bar Plot with Churn Grouped") +
  scale_fill_manual(values = c("#142850", "lightblue"))

# Internet Service Bar Plot with Churn Grouped
g2 <- ggplot(data = df, aes(x = InternetService)) +
  geom_bar(aes(fill = Churn), position = "dodge") +
  labs(x = "Internet Service", y = "Count", title = "Internet Service Bar Plot with Churn Grouped") +
  scale_fill_manual(values = c("#142850", "lightblue"))

# Senior Citizen Bar Plot with Churn Grouped
g3 <- ggplot(data = df, aes(x = SeniorCitizen)) +
  geom_bar(aes(fill = Churn), position = "dodge") +
  labs(x = "Senior Citizen", y = "Count", title = "Senior Citizen Distribution Bar Plot with Churn Grouped") +
  scale_fill_manual(values = c("#142850", "lightblue"))

# Gender Bar Plot with Churn Grouped
g4 <- ggplot(data = df, aes(x = gender)) +
  geom_bar(aes(fill = Churn), position = "dodge") +
  labs(x = "Gender", y = "Count", title = "Gender Distribution Bar Plot with Churn Grouped") +
  scale_fill_manual(values = c("#142850", "lightblue"))

g1
g2
g3
g4
```

## 4) Correlation

```{r}
df_num <- df[, sapply(df, is.numeric)]
print(cor((df_num)))
```

```{r}
library(reshape2)
corr_melted <- melt(cor(df_num))
corr_melted_sorted <- corr_melted %>% arrange(desc(value))
head(corr_melted_sorted, 10)
```

```{r}
library(ggplot2)
ggplot(corr_melted, aes(x = Var1, y = Var2, fill = value)) +
geom_tile() +
scale_fill_gradient2(low = "#DEFCF9", mid = "#3282B8", high = "#142850",
name = "Pearson\nCorrelation") + theme(axis.text.x = element_text(angle = 90, vjust = 0.7, hjust = 1))
```

# Predictive Modelling

```{r}
library(caret)

# Convert Churn variable to factor
df$Churn <- as.factor(df$Churn)

# Perform oversampling
dfb <- upSample(x = df[, -which(names(df) == "Churn")],
                        y = df$Churn,
                        yname = "Churn")

# Check the balanced class distribution
table(dfb$Churn)

```

```{r}
# Before oversampling
b1 = ggplot(data = df, aes(x = Churn)) +
  geom_bar(fill = c("#142850", "lightblue")) +
  labs(x = "Churn", y = "Count", title = "Before Oversampling")

# Create a bar plot for the distribution of Churn in balanced_df
b2 = ggplot(data = dfb, aes(x = Churn)) +
  geom_bar(fill = c("#142850", "lightblue")) +
  labs(x = "Churn", y = "Count", title = "After Oversampling")

library(gridExtra)
grid.arrange(b1, b2, ncol=2)
```

```{r}
# factor
dfb$gender <- factor(dfb$gender)
dfb$SeniorCitizen <- factor(dfb$SeniorCitizen)
dfb$Partner <- factor(dfb$Partner)
dfb$Dependents <- factor(dfb$Dependents)
dfb$PhoneService <- factor(dfb$PhoneService)
dfb$MultipleLines <- factor(dfb$MultipleLines)
dfb$InternetService <- factor(dfb$InternetService)
dfb$OnlineSecurity <- factor(dfb$OnlineSecurity)
dfb$OnlineBackup <- factor(dfb$OnlineBackup)
dfb$DeviceProtection <- factor(dfb$DeviceProtection)
dfb$TechSupport <- factor(dfb$TechSupport)
dfb$StreamingTV <- factor(dfb$StreamingTV)
dfb$StreamingMovies <- factor(dfb$StreamingMovies)
dfb$Contract <- factor(dfb$Contract)
dfb$PaperlessBilling <- factor(dfb$PaperlessBilling)
dfb$PaymentMethod <- factor(dfb$PaymentMethod)
dfb$Churn <- factor(dfb$Churn)
```

```{r}
head(dfb)
```

```{r}
# split the data into training and testing sets
library(caTools)
set.seed(1)
split <- sample.split(dfb$Churn, SplitRatio = 0.8)
train <- subset(dfb, split == TRUE)
test <- subset(dfb, split == FALSE)
```

```{r}
n <- names(train)
excluded_vars <- c("Churn")
f <- as.formula(paste("Churn ~", paste(n[!n %in% excluded_vars], collapse = " + ")))
```

```{r}
library(randomForest)
modelRF <- randomForest(f,data=train)
print(modelRF)
```

```{r}
#create a data frame to compare the predicted value with the ground truth
predictedRF <- predict(modelRF,test)
```

```{r}
evaluation = confusionMatrix(predictedRF, test$Churn)
evaluation
```

```{r}
importance(modelRF)
```

```{r}
varImpPlot(modelRF)
```

```{r}
importance_df <- as.data.frame(importance(modelRF))
importance_df$Variables <- rownames(importance_df)

ggplot(importance_df, aes(x = Variables, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Feature Importance",
       x = "Variables",
       y = "MeanDecreaseGini")
```
